Pulling oral and IV data from CvTdb

### 2025-05-20 data-raw/pulling_oral_iv_data_cvtdb.Rmd
Updating data-raw with changes to vignettes and built-in model function arguments.

```{r}
library(DBI)
library(dplyr)
library(tidyr)
```

@@

quick_stats <- function(.data) {
  stopifnot(is.data.frame(.data))
  cli::cli_h2("Statistics on the current data")
  message(paste0("Total number of data points: ", nrow(.data)))
  message("Unique Species and data counts:")
  print(table(.data$species))

  message(paste0("Number of unique studies: ",
                 length(unique(.data$fk_study_id))))
  message(paste0("Number of unique series: ",
                 length(unique(.data$fk_series_id))))
  message(paste0("Number of unique chemicals: ",
                 length(unique(.data$analyte_dtxsid))))
  message(paste0("Unique Dose units: ",
                 toString(unique(.data$dose_level_units_original))))
  message(paste0("Unique Concentration units: ", 
                 toString(unique(.data$conc_units_original))))
  message("Data by unique route of administration:")
  print(table(.data$administration_route_normalized))
  message("Data by medium collected:")
  print(table(.data$conc_medium_normalized))
  message("Data by curation set:")
  print(table(.data$curation_set_tag))
}

@@

```{r diff_log_function}
# Now I want to write a quick function that will log relevant changes
# from previous 'cvt' objects
devtools::load_all() # So this loads the previous cvt object
diff_log_cvt <- function(new_cvt,
                         key_id = "conc_time_id",
                         grp_ids = c("analyte_dtxsid",
                                     "species"),
                         notes = character(0L)) {
  stopifnot(
    key_id %in% colnames(cvt),
    key_id %in% colnames(new_cvt),
    all(grp_ids %in% colnames(cvt)),
    all(grp_ids %in% colnames(new_cvt))
  )
  if (length(unique(new_cvt[[key_id]])) != NROW(new_cvt)) {
    stop("key_id is not not a column that uniquely identifies this data.")
  }

@@

```{r SQL_queries, warning=FALSE}
# Need to make this in a specified ordered
# First all IDs (so I can check them quickly)
# Note that this uses an ALIAS for each table which is later specified in
# the FROM and LEFT JOIN statments
query <- paste0(
  "SELECT distinct ",
  # Conc_Time_Values table fields
  "a.id, a.fk_series_id, a.time_original, a.time_hr, ",
  "a.conc_original, a.conc_sd_original, a.conc, a.conc_sd, ",

@@

  "c.dose_volume, c.dose_volume_units, c.dose_vehicle, ",
  "c.dose_duration, c.dose_duration_units, ",
  ## dose_frequency dictionary fields
  "j.dose_frequency_original, j.dose_frequency_normalized, ",
  "c.fasting_period, ",
  "c.dose_level_normalized, c.dose_level_original, c.dose_level_units_original, ",
  ## conc_medium dictionary fields
  "i.conc_medium_original, i.conc_medium_normalized, ", 
  ## administration_route dictionary fields
  "h.administration_route_original, h.administration_route_normalized, ",
  ## administration_method dictionary fields
  "g.administration_method_original, g.administration_method_normalized, ",
  ## administration_form dictionary fields
  "f.administration_form_original, f.administration_form_normalized, ",

@@

  # Join to dictionary tables
  "LEFT JOIN administration_form_dict f ON c.fk_administration_form_id = f.id ",
  "LEFT JOIN administration_method_dict g ON c.fk_administration_method_id = g.id ",
  "LEFT JOIN administration_route_dict h ON c.fk_administration_route_id = h.id ",
  "LEFT JOIN conc_medium_dict i ON b.fk_conc_medium_id = i.id ",
  "LEFT JOIN dose_frequency_dict j ON c.fk_dose_frequency_id = j.id ",

  # Rename chemical fields for dosed vs. analyzed chemical record foreign keys
  "LEFT JOIN (SELECT id, dsstox_substance_id as dosed_chem_dtxsid, ",
  "chemical_name_original as dosed_chem_name_original, ", 
  "dsstox_casrn as dosed_chem_casrn, preferred_name as dosed_chem_name ",
  "FROM chemicals) as k ON c.fk_dosed_chemical_id = k.id ",
  "LEFT JOIN (SELECT id, dsstox_substance_id as analyzed_chem_dtxsid, ",
  "chemical_name_original as analyzed_chem_name_original, ",
  "dsstox_casrn as analyzed_chem_casrn, preferred_name as analyzed_chem_name ",
  "FROM chemicals) as l ON b.fk_analyzed_chemical_id = l.id ",

  # Filtering steps
  "WHERE h.administration_route_normalized IN ('iv', 'oral') AND ",
  "(i.conc_medium_normalized IN ('blood', 'plasma', 'excreta') OR ",
  "i.conc_medium_original IN ('blood', 'plasma', 'excreta'))"
)

@@

# Need to change column names for this
{
  names(sql_query)[1] <- "conc_time_id"
}

# Unfortunately, there are some "fancy" or curly quotes that need to be replaced
# with "straight" quotes to be more machine read-able
force_straight_quotes <- function(x) {
  temp <- gsub("(’|′)", "'", x)
  temp <- gsub("(, | ,)", ",", temp)
  return(temp)
}
force_straight_quotes("3,3’,4,4’-tetrachloro-5-biphenylol")

# Harmonize values for easier analysis
cvt_df <- sql_query %>%
  mutate(
    across(ends_with("id"), as.character),
    # Note that sometimes fk_test_chemical_id is NA
    across(all_of(c("time_original",
                    "time_hr",
                    "conc_original",
                    "conc_sd_original",
                    "conc",
                    "conc_sd",
                    "dose_level_original",
                    "dose_level_normalized",
                    "loq")), as.numeric),
    # conc, conc_sd_original, and conc_original
    # have NA/NE/ND values that are coerced to NAs
    year = as.character(year),
    species = tolower(species),
    sex = tolower(sex),
    analyte_name_original = force_straight_quotes(tolower(analyte_name_original)),
    analyte_name_secondary_original = force_straight_quotes(tolower(
      analyte_name_secondary_original
      )),
    dose_volume_units = stringr::str_extract(dose_volume,
                                    pattern = "(?<=[:space:]).+"),
    dose_volume = as.numeric(stringr::str_extract(dose_volume,
                                         pattern = "^(\\d|\\.)*")),
    dose_duration_units = stringr::str_extract(dose_duration,
                                      pattern = "(?<=[space]).+"),
    dose_duration = as.numeric(stringr::str_extract(dose_duration,
                                pattern = "^\\d")),
    radiolabeled = ifelse(is.na(radiolabeled), FALSE, as.logical(radiolabeled)),
    n_subjects_normalized = ifelse(n_subjects_in_series %in% "NR", NA_real_,
      as.numeric(stringr::str_extract(n_subjects_in_series,
                                                 pattern = "\\d+")))
  )

cvt_df <- cvt_df %>%
  mutate(conc_medium_normalized = ifelse(is.na(conc_medium_normalized),
                                         conc_medium_original,
                                         conc_medium_normalized))

# There are some missing DTXSIDs that we need to fix
cvt_df %>% 
    filter(is.na(analyte_dtxsid),
           !(radiolabeled %in% TRUE)) %>%
  distinct(fk_analyzed_chemical_id,fk_dosed_chemical_id,
           analyte_dtxsid, analyte_name_original,
           analyte_name_secondary_original) %>%
  View("NA_dtxsid")

dtxsid_query <- paste0(
  "SELECT distinct * FROM chemicals"
)

dtxsid_query <- query_db(query = dtxsid_query,
         schema = "cvt")

dtxsid_query <- dtxsid_query %>%
  filter(id %in% as.numeric(
    union(cvt_df$fk_analyzed_chemical_id, cvt_df$fk_dosed_chemical_id))
  ) %>%
  mutate(id = as.character(id),
         chemical_name_original = force_straight_quotes(tolower(
           chemical_name_original
         )),
         chemical_name_secondary_original = force_straight_quotes(tolower(
           chemical_name_secondary_original
         ))
  )

cvt_df %>% 
  filter(fk_analyzed_chemical_id != fk_dosed_chemical_id,
         !(radiolabeled %in% TRUE)) %>%
  distinct(fk_analyzed_chemical_id,fk_dosed_chemical_id,fk_test_chemical_id) %>%
  left_join(dtxsid_query %>%
              select("fk_analyzed_chemical_id" = "id",
                     "analyzed_chemical_name" = "chemical_name_original")) %>%
  left_join(dtxsid_query %>%
              select("fk_dosed_chemical_id" = "id",
                     "dosed_chemical_name" = "chemical_name_original")) %>%
  View("errors")

# Rather than hope most of these chemicals have dtxsids within CvTdb,
# I cross-referenced it with the CompTox database batch search.
dtxsid_batch_idless <- readr::read_csv(
  "data-raw/CCD-Batch-Search_2025-04-17_DTXSIDless.csv",
  na = "N/A"
) %>% filter(!is.na(DTXSID)) %>%
  select("dsstox_substance_id" = "DTXSID",
         "chemical_name_original" = "INPUT") %>%
  distinct()

quick_message(c(
  "Erroneous DTXSID assignments (identified via duplication):",
  "perfluorohexanesulfonate == DTXSID80873012",
  "tacrolimus == DTXSID5046354",
  "hippuric acid == DTXSID9046073"
))

dtxsid_batch_idless <- dtxsid_batch_idless %>%
  mutate(dsstox_substance_id = case_when(
    chemical_name_original == "perfluorohexanesulfonate" ~ "DTXSID80873012",
    chemical_name_original == "tacrolimus" ~ "DTXSID5046354",
    chemical_name_original == "hippuric acid" ~ "DTXSID9046073",
    .default = dsstox_substance_id
  )) %>% distinct()

# Fixing the missing DTXSID to prevent premature filtering
cvt_df <- cvt_df %>%
  left_join(
    dtxsid_batch_idless,
    by = join_by(analyte_name_original == chemical_name_original)
  ) %>%
  mutate(analyte_dtxsid = coalesce(analyte_dtxsid, dsstox_substance_id)) %>%
  select(!dsstox_substance_id)

@@

# Now we need to pre-emptively address the chemicals that have different
# dosed_id and analyzed_ids but are actually the same chemical
quick_message(c(
  "The following have resolvable errors:",
  "If dosed chemical id is 47, change this to 46 to match analyte id. tamoxifen, salt dosed.",
  "If dosed chemical id is 893, change this to 929 to match analyte id. trichloroethylene, duplicate.",
  "If dosed chemical id is 671, change to 672 to match analyte id. PFHS, salt dosed.",
  "If dosed chemical id is 673, change to 675 to match analyte id. PFNA, capitalization.",
  "If dosed chemical id is 674, change to 676 to match analyte id, PFDA, capitalization."
))

cvt_df <- cvt_df %>%
  rowwise() %>%
  mutate(fk_dosed_chemical_id = case_when(
    fk_dosed_chemical_id == "47" &&
      fk_analyzed_chemical_id == "46" ~ "46",
    fk_dosed_chemical_id == "893" &&
      fk_analyzed_chemical_id == "929" ~ "929",
    fk_dosed_chemical_id == "671" &&
      fk_analyzed_chemical_id == "672" ~ "672",
    fk_dosed_chemical_id == "673" &&
      fk_analyzed_chemical_id == "673" ~ "673",
    fk_dosed_chemical_id == "674" &&
      fk_analyzed_chemical_id == "676" ~ "676",
    .default = fk_dosed_chemical_id
  )) %>%
  ungroup()

# glimpse(dq_cols)

# names(cvt_df)

# study_query <- query_db(query = "SELECT distinct * FROM studies",
#          schema = "cvt")

@@

cat("Current # chemicals: ", length(unique(cvt_df$analyte_dtxsid)), "\n")
cat("Number of chemicals where analyte_dtxsid is NA: ",
    sum(is.na(unique(cvt_df[c("fk_analyzed_chemical_id",
                               "analyte_dtxsid")])$analyte_dtxsid)),
    "\n")

cat("Number of radiolabelling experiments: ",
    sum(distinct(cvt_df, analyte_dtxsid, radiolabeled) %>% pull(radiolabeled) %>%
          {. %in% TRUE}
        )
    )
distinct(cvt_df, analyte_dtxsid, analyte_name_original, radiolabeled) %>%
  filter(radiolabeled %in% TRUE, !is.na(analyte_dtxsid))

cat("Number of chemicals with non-matching ids ",
    cvt_df %>%
      filter(!is.na(analyte_dtxsid),
             !(radiolabeled %in% TRUE),
             fk_analyzed_chemical_id != fk_dosed_chemical_id) %>%
      distinct(analyte_dtxsid) %>%
      NROW()
)

cvt_df <- cvt_df %>%
  filter(!(radiolabeled %in% TRUE),
         !is.na(analyte_dtxsid),
         (fk_analyzed_chemical_id == fk_dosed_chemical_id)) %>%
  distinct()

quick_message(text = c("Filtering out:",
                       "Missing {.code analyte_dtxsid}",
                       "Radiolabeling experiments",
                       "Data where the analyzed chemicals and dosed chemicals are different"))
quick_stats(cvt_df)
diff_log_cvt(cvt_df)
```

@@

```{r preliminary_filtering}
# Dose Units: Need mg/kg... = ug/mg = ng/ug, and kg = L
# Conc Units: Need mg/L... = ug/mL = ng/uL = pg/nL, and kg = L, g = mL
cvt_df <- cvt_df %>% mutate(
  invivPK_dose_level_units = "mg/kg",
  invivPK_conc_units = "ug/mL",
  invivPK_conc = NA_real_,
  invivPK_dose_level = NA_real_
  )
# Set the equivalent units
equiv_mgkg <- c("mg/kg", "ug/g", "ng/mg")
equiv_mgL <- c("mg/L", "ug/mL", "ng/uL", "mg/kg", "mg/g", "ug/mg")

>>>>

```{r}
library(DBI)
devtools::load_all() # So this loads the previous cvt object
library(dplyr)
library(tidyr)
library(readr)
```

@@

quick_stats <- function(.data, old_version = FALSE) {
  stopifnot(is.data.frame(.data))

  cli::cli_h2("Statistics on the current data")
  message(paste0("Total number of data points: ", nrow(.data)))
  message("Unique Species and data counts:")
  print(table(.data$species))

  message("Number of unique studies: ",
          length(unique(.data$fk_study_id)))
  message("Number of unique series: ",
          length(unique(.data$fk_series_id)))

  if (old_version) {
    message("Number of unique chemicals: ",
          length(unique(.data$analyte_dtxsid)),
          "\n")
    chem_by_medium <- .data[c("conc_medium_normalized", "analyte_dtxsid")] %>%
      filter(!is.na(analyte_dtxsid)) %>%
      distinct()
    chem_by_curset <- .data[c("curation_set_tag", "analyte_dtxsid")] %>%
      filter(!is.na(analyte_dtxsid)) %>%
      distinct()
  } else {
    message("Number of unique chemicals: ",
          length(unique(.data$analyzed_chem_dtxsid)),
          "\n")
    chem_by_medium <- .data[c("conc_medium_normalized", "analyzed_chem_dtxsid")] %>%
      filter(!is.na(analyzed_chem_dtxsid)) %>%
      distinct()
    chem_by_curset <- .data[c("curation_set_tag", "analyzed_chem_dtxsid")] %>%
      filter(!is.na(analyzed_chem_dtxsid)) %>%
      distinct()
  }

  message("Unique Dose units: ",
          toString(unique(.data$dose_level_units_original)),
          "\n")
  if (all(is.na(.data$conc_units_normalized))) {
    message("Unique Concentration units: ", 
            toString(unique(.data$conc_units_original)),
            "\n")
  } else {
    message("Unique Concentration units: ", 
            toString(unique(.data$conc_units_normalized)),
            "\n")
  }
  message("Data by unique route of administration:")
  print(table(.data$administration_route_normalized))
  message("Data by medium collected:")
  print(table(.data$conc_medium_normalized))
  message("Data by curation set:")
  print(table(.data$curation_set_tag))
  message("\n\nUnique chemicals per medium:")
  print(table(chem_by_medium$conc_medium_normalized))
  message("Unique chemicals per curation set:")
  print(table(chem_by_curset$curation_set_tag))
}

@@

```{r diff_log_function}
# Now I want to write a quick function that will log relevant changes
# from previous 'cvt' objects
diff_log_cvt <- function(new_cvt,
                         key_id = "conc_time_id",
                         grp_ids = c("analyzed_chem_dtxsid",
                                     "species"),
                         notes = character(0L)) {
  stopifnot(
    key_id %in% colnames(cvt),
    key_id %in% colnames(new_cvt),
    all(grp_ids %in% colnames(cvt)),
    all(grp_ids %in% colnames(new_cvt))
  )
  if (length(unique(new_cvt[[key_id]])) != NROW(new_cvt)) {
    stop("key_id is not not a column that uniquely identifies this data.")
  }

@@

```{r SQL_queries, warning=FALSE}
# Need to make this in a specified ordered
# First all IDs (so I can check them quickly)
# Note that this uses an ALIAS for each table which is later specified in
# the FROM and LEFT JOIN statements
series_chems <- paste0(
  "SELECT DISTINCT a.id, a.fk_analyzed_chemical_id, ",
  "b.analyzed_chem_dtxsid, b.analyzed_chem_name_original, ",
  "b.analyzed_chem_name_secondary_original, ",
  "b.analyzed_chem_casrn, b.analyzed_chem_name, ",
  "a.time_units_original, a.conc_units_original, ",
  "a.conc_units_normalized, a.loq, a.loq_units, ",
  "a.n_subjects_in_series, a.radiolabeled, a.fk_study_id, ",
  "a.qc_notes, a.qc_status, a.qc_flags, c.conc_medium_normalized ",
  "FROM series a ",
  "LEFT JOIN (SELECT id, dsstox_substance_id as analyzed_chem_dtxsid, ",
  "chemical_name_original as analyzed_chem_name_original, ",
  "chemical_name_secondary_original as analyzed_chem_name_secondary_original, ",
  "dsstox_casrn as analyzed_chem_casrn, preferred_name as analyzed_chem_name ",
  "FROM chemicals) as b ON a.fk_analyzed_chemical_id = b.id ",
  "LEFT JOIN conc_medium_dict as c ON a.fk_conc_medium_id = c.id"
)

study_docs <- paste0(
  "SELECT a.id, a.fk_extraction_document_id, a.fk_dosed_chemical_id, ",
  "b.pmid, b.doi, b.year, b.other_study_identifier, ",
  "b.url, b.extracted, b.curation_set_tag ",
  "FROM studies a ",
  "LEFT JOIN documents b ON a.fk_extraction_document_id = b.id"
)

audit_query <- query_db(paste0("SELECT DISTINCT * FROM cvt_audit ",
                               "WHERE fk_table_name IN ('documents', 'series', 'studies')"),
                        "cvt")
study_query <- query_db(study_docs, "cvt")
series_query <- query_db(series_chems, "cvt")
dtxsid_query <- query_db("SELECT DISTINCT * FROM chemicals", "cvt") %>%
  mutate(
    across(ends_with("id"), as.character),
    across(contains("chemical_name"), \(x) {
      tolower(force_ascii_quotes(x))
    })
  )

# There are a lot of duplicate DTXSIDs
query <- paste0(
  "SELECT distinct ",
  # Conc_Time_Values table fields
  "a.id, a.fk_series_id, a.time_original, a.time_hr, ",
  "a.conc_original, a.conc_sd_original, a.conc, a.conc_sd, ",

@@

  "c.dose_volume, c.dose_volume_units, c.dose_vehicle, ",
  "c.dose_duration, c.dose_duration_units, ",
  ## dose_frequency dictionary fields
  "j.dose_frequency_original, j.dose_frequency_normalized, ",
  "c.fasting_period, ",
  "c.dose_level_normalized, c.dose_level_original, c.dose_level_units_original, ",
  ## conc_medium dictionary fields
  "i.conc_medium_original, i.conc_medium_normalized, ", 
  ## administration_route dictionary fields
  "h.administration_route_original, h.administration_route_normalized, ",
  ## administration_method dictionary fields
  # "g.administration_method_original, g.administration_method_normalized, ",
  ## administration_form dictionary fields
  # "f.administration_form_original, f.administration_form_normalized, ",

@@

  # # Join to dictionary tables
  # "LEFT JOIN administration_form_dict f ON c.fk_administration_form_id = f.id ",
  # "LEFT JOIN administration_method_dict g ON c.fk_administration_method_id = g.id ",
  "LEFT JOIN administration_route_dict h ON c.fk_administration_route_id = h.id ",
  "LEFT JOIN conc_medium_dict i ON b.fk_conc_medium_id = i.id ",
  "LEFT JOIN dose_frequency_dict j ON c.fk_dose_frequency_id = j.id ",

  # Rename chemical fields for dosed vs. analyzed chemical record foreign keys
  "LEFT JOIN (SELECT id, dsstox_substance_id as dosed_chem_dtxsid, ",
  "chemical_name_original as dosed_chem_name_original, ", 
  "chemical_name_secondary_original as dosed_chem_name_secondary_original, ",
  "dsstox_casrn as dosed_chem_casrn, preferred_name as dosed_chem_name ",
  "FROM chemicals) as k ON c.fk_dosed_chemical_id = k.id ",
  "LEFT JOIN (SELECT id, dsstox_substance_id as analyzed_chem_dtxsid, ",
  "chemical_name_original as analyzed_chem_name_original, ",
  "chemical_name_secondary_original as analyzed_chem_name_secondary_original, ",
  "dsstox_casrn as analyzed_chem_casrn, preferred_name as analyzed_chem_name ",
  "FROM chemicals) as l ON b.fk_analyzed_chemical_id = l.id ",

  # Filtering steps
  "WHERE h.administration_route_normalized IN ('iv', 'oral') AND ",
  "(i.conc_medium_normalized IN ('blood', 'plasma') OR ",
  "i.conc_medium_original IN ('blood', 'plasma')) AND ",
  "(j.dose_frequency_normalized IN ('Once') OR e.curation_set_tag IN ('CVT_Showa'))"
)

@@

# Need to change the column name for this
names(sql_query)[1] <- "conc_time_id"

# cvtdb_original <- sql_query
# save(cvtdb_original, file = "data/cvtdb_original.rda", compress = "bzip2")
```

```{r harmonizing}
load("data/cvtdb_original.rda")
sql_query <- cvtdb_original
rm(cvtdb_original)

# Harmonize values for easier analysis
cvt_df <- sql_query %>%
  mutate(
    across(ends_with("id"), as.character),
    # Note that sometimes fk_test_chemical_id is NA
    across(all_of(c("time_original",
                    "time_hr",
                    "conc_original",
                    "conc_sd_original",
                    "conc",
                    "conc_sd",
                    "dose_level_original",
                    "dose_level_normalized",
                    "loq")), as.numeric),
    # conc, conc_sd_original, and conc_original
    # have NA/NE/ND values that are coerced to NAs
    year = as.character(year),
    species = tolower(species),
    sex = tolower(sex),
    across(contains("chem_name"), \(x) {
      tolower(force_ascii_quotes(x))
    }),
    dose_volume_units = ifelse(
      dose_volume_units %in% c("NR", "ND"),
      NA_character_,
      gsub(" / ", "/", gsub("ml", "mL", dose_volume_units))
      ),
    dose_volume = signif(as.numeric(dose_volume), digits = 7),
    dose_duration_units = trimws(dose_duration_units),
    dose_duration = as.numeric(stringr::str_extract(dose_duration,
                                pattern = "^\\d+")),
    dose_frequency_normalized = "Once", # Showa is not annotated
    radiolabeled = ifelse(is.na(radiolabeled), FALSE, as.logical(radiolabeled)),
    # for n_subject ranges, take integer value of median 
    # otherwise pull number or NA
    n_subjects_normalized = case_when(
      n_subjects_in_series %in% "NR" ~ NA_real_,
      grepl("-", n_subjects_in_series) ~ floor(mean(
        as.numeric(unlist(strsplit(n_subjects_in_series, split = "-"))),
        na.rm = TRUE
        )),
      .default = as.numeric(
        stringr::str_extract(n_subjects_in_series, pattern = "\\d+")
        )
      )
  )

## Some data malformations

There are some duplications in chemical entries such that not all unique chemical ids
represent a unique DTXSID, but rather a unique original name entry.
To fix this, when DTXSIDs for dosed and analyzed chemicals match, I will
be setting `fk_dosed_chemical_id` to `fk_analyzed_chemical_id`.

There is also some radiolabeling experiments that aren't described as such.
We can use `stringr::str_detect(x,"(13|14)c")` to recode `radiolabeled = TRUE`
for these.

Then we should address the fact that the `invivoPKfit_2.0.0` version of the `cvt`
object has 228 chemicals, while the recent query only has 145. This is a
difference of 83.

By hand, I can add back the DTXSIDs for about 15 chemicals. Some of these will
be ChEMBLIDs actually.

```{r malformed_data}
# First quick check to see if any chemical_ids are NA
any(is.na(cvt_df$fk_analyzed_chemical_id))
any(is.na(cvt_df$fk_dosed_chemical_id))
length(unique(cvt_df$analyzed_chem_dtxsid))

cvt_df <- cvt_df %>%
  rowwise() %>%
  mutate(fk_dosed_chemical_id = ifelse(
    identical(analyzed_chem_dtxsid, dosed_chem_dtxsid),
    fk_analyzed_chemical_id,
    fk_dosed_chemical_id
  )) %>%
  ungroup()

cvt_df %>%
  distinct(analyzed_chem_name_original, dosed_chem_name_original, radiolabeled) %>%
  filter(!radiolabeled) %>%
  filter(stringr::str_detect(dosed_chem_name_original, "(13|14)c"))

cvt_df <- cvt_df %>%
  mutate(
    radiolabeled = ifelse(
      stringr::str_detect(dosed_chem_name_original, "(13|14)c"),
      TRUE,
      radiolabeled
    )
  )

quick_stats(cvt_df)
quick_stats(cvt_2.0.0, old_version = TRUE)
# Note unique chemical counts

# Rather than hope most of these chemicals have dtxsids within CvTdb,
# I cross-referenced it with the CompTox database batch search.
# Here I have curated the input by hand
dtxsid_byhand_idless <- data.frame(
  chem_name_original = c(
    "paracetmol", #misspelling
    "antipyrene", #synonym phenazone
    "r-(+)-metoprolol", # using R/S and +/- confused search
    "s-(-)-metoprolol",
    "deacetyl diltiazem", # should be desacetyl
    "valproate (vpa)", # valproic acid
    "valproic acid",
    "clotiapine", # clothiapine
    "padsevonil", # found a ChEMBL_ID...
    "2,2',4,4',5,'5-hexachlorobiphenyl",
    "2,4,5,2',4','5-hexachlorobiphenyl", # number/prime order confused
    "quinidine sulfate",
    "quinidine",
    "quinidine gluconate",
    "furosemide",
    "theophylline",
    "fentanyl",
    "ammonium perfluorooctonoate",
    "perfluorohexanesulfonate anion", # anion confused search
    "3'-azido-3'-deoxythymidine (aids)", # parenthetical confuses search
    "3'-azido-3'-deoxythymidine", # repeat/duplicate of above
    "2,4,5,2'5'-pentachlorobiphenyl", # number/prime order incorrect
    "4,4'-sufonylbisphenol-d8", # deuterium but same as bisphenol S
    "hydroxy-tebuconazole", # CAS 212267-64-6
    "3'-amino-3'-deoxythimidine",
    "bosentan sodium salt",
    "bosentan micronized free sulfonamide",
    "dichloroacetate",
    "fluorotelomer acid 7+3",
    "furosemide",
    "morphine hydrochloride",
    "perfluorohexanesulfonate,potassium salt",
    "potassium perfluorooctane",
    "sodium hexobarbital"
  ),
  comptox_dtxsid = c(
    "DTXSID2020006",
    "DTXSID6021117",
    "DTXSID20230862",
    "DTXSID20230861",
    "DTXSID00881093",
    "DTXSID6023733",
    "DTXSID6023733",
    "DTXSID5022851",
    "CHEMBL4297521",
    "DTXSID2032180",
    "DTXSID2032180",
    "DTXSID101017271",
    "DTXSID4023549",
    "DTXSID80102749",
    "DTXSID6020648",
    "DTXSID5021336",
    "DTXSID9023049",
    "DTXSID8037708",
    "DTXSID80873012",
    "DTXSID8020127",
    "DTXSID8020127",
    "DTXSID8038304",
    "DTXSID3022409",
    "DTXSID50891627",
    "DTXSID00966879",
    "DTXSID7046627",
    "DTXSID7046627",
    "DTXSID2020428",
    "DTXSID90382620",
    "DTXSID6020648",
    "DTXSID9023336",
    "DTXSID3037709",
    "DTXSID0059794",
    "DTXSID9023122"
  )
) %>%
  distinct()

# Fill in the missing DTXSIDs for these
cvt_df <- cvt_df %>%
  # Fill for analyzed chemicals
  left_join(
    dtxsid_byhand_idless %>%
      inner_join(cvt_df %>%
                   distinct(analyzed_chem_name_original, fk_analyzed_chemical_id),
                 by = join_by(chem_name_original == analyzed_chem_name_original)),
    by = join_by(fk_analyzed_chemical_id,
                 analyzed_chem_name_original == chem_name_original)
  ) %>%
  mutate(analyzed_chem_dtxsid = coalesce(analyzed_chem_dtxsid, comptox_dtxsid)) %>%
  select(!comptox_dtxsid) %>%
  # Now do the same for dosed chemical
  left_join(
    dtxsid_byhand_idless %>%
      inner_join(
        cvt_df %>%
          distinct(dosed_chem_name_original, fk_dosed_chemical_id),
        by = join_by(chem_name_original == dosed_chem_name_original)
      ),
    by = join_by(fk_dosed_chemical_id,
                 dosed_chem_name_original == chem_name_original)
  ) %>%
  mutate(dosed_chem_dtxsid = coalesce(dosed_chem_dtxsid, comptox_dtxsid)) %>%
  select(!comptox_dtxsid)

# Now that we've added some missing DTXSIDs, redo the chemical id 
cvt_df <- cvt_df %>%
  rowwise() %>%
  mutate(fk_dosed_chemical_id = ifelse(
    identical(analyzed_chem_dtxsid, dosed_chem_dtxsid),
    fk_analyzed_chemical_id,
    fk_dosed_chemical_id
  )) %>%
  ungroup()

any(is.na(cvt_df$fk_analyzed_chemical_id))
any(is.na(cvt_df$fk_dosed_chemical_id))

@@

# Now we need to pre-emptively address the chemicals that have different
# dosed_id and analyzed_ids but are actually the same chemical
cvt_df %>%
  filter(fk_analyzed_chemical_id != fk_dosed_chemical_id, 
         analyzed_chem_dtxsid != dosed_chem_dtxsid) %>%
  distinct(analyzed_chem_name_original, dosed_chem_name_original,
           fk_analyzed_chemical_id, fk_dosed_chemical_id,
           analyzed_chem_dtxsid, dosed_chem_dtxsid) %>%
  arrange(analyzed_chem_name_original) %>%
  head()
  # View()

# TO FIX: SALTS
# perfluorooctanesulfonate 692 -> 694, 763 -> 764
# perfluorooctanoate 761 -> 762
# phenobarbital 1147 -> 291
# perfluorobutyrate (two different salts, could be interesting but removing for now)
# perfluorobutanesulfonate 679 -> 680, 777 -> 778
# perfluorohexanesulfonate 671 -> 672, 791 -> 792, 681 -> 682
# tamoxifen 1008 -> 1010
# thiopenal 1148 -> 1150
# imipramine 1158 -> 1159, 1154 -> 1159
# valproate 1163 -> 1164
# nitrite 1013 -> 1014
# odansetron 1117 -> 1118

cvt_df <- cvt_df %>%
  mutate(
    fk_dosed_chemical_id = case_when(
      fk_dosed_chemical_id %in% "692" &
        fk_analyzed_chemical_id %in% "694" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "763" &
        fk_analyzed_chemical_id %in% "764" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "761" &
        fk_analyzed_chemical_id %in% "762" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "1147" &
        fk_analyzed_chemical_id %in% "291" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "679" &
        fk_analyzed_chemical_id %in% "680" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "777" &
        fk_analyzed_chemical_id %in% "778" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "671" &
        fk_analyzed_chemical_id %in% "672" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "791" &
        fk_analyzed_chemical_id %in% "792" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "681" &
        fk_analyzed_chemical_id %in% "682" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "1008" &
        fk_analyzed_chemical_id %in% "1010" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "1148" &
        fk_analyzed_chemical_id %in% "1150" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "1158" &
        fk_analyzed_chemical_id %in% "1159" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "1154" &
        fk_analyzed_chemical_id %in% "1159" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "1163" &
        fk_analyzed_chemical_id %in% "1164" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "1013" &
        fk_analyzed_chemical_id %in% "1014" ~ fk_analyzed_chemical_id,
      fk_dosed_chemical_id %in% "1117" &
        fk_analyzed_chemical_id %in% "1118" ~ fk_analyzed_chemical_id,
      .default = fk_dosed_chemical_id
    )
  )

@@

cat("Current # chemicals: ", length(unique(cvt_df$analyzed_chem_dtxsid)), "\n")
cat("Number of chemicals where analyzed_chem_dtxsid is NA: ",
    sum(is.na(unique(cvt_df[c("fk_analyzed_chemical_id",
                               "analyzed_chem_dtxsid")])$analyzed_chem_dtxsid)),
    "\n")

cat("Number of radiolabelling experiments: ",
    sum(distinct(cvt_df, analyzed_chem_dtxsid, radiolabeled) %>%
          pull(radiolabeled) %>%
          {. %in% TRUE}
        )
    )
distinct(cvt_df, analyzed_chem_dtxsid,
         analyzed_chem_name_original, radiolabeled) %>%
  filter(radiolabeled %in% TRUE, !is.na(analyzed_chem_dtxsid))

cat("Number of chemicals with non-matching ids ",
    cvt_df %>%
      filter(!is.na(analyzed_chem_dtxsid),
             !(radiolabeled %in% TRUE),
             fk_analyzed_chemical_id != fk_dosed_chemical_id) %>%
      distinct(analyzed_chem_dtxsid) %>%
      NROW()
)

quick_stats(cvt_df)
# This should be the base CvTdb prior to filtering steps

@@

```{r preliminary_filtering}

cvt_df <- cvt_df %>%
  filter(!(radiolabeled %in% TRUE),
         !is.na(analyzed_chem_dtxsid),
         (fk_analyzed_chemical_id == fk_dosed_chemical_id)) %>%
  distinct()

quick_message(text = c("Filtering out:",
                       "Missing {.code analyzed_chem_dtxsid}",
                       "Radiolabeling experiments",
                       "Data where the analyzed chemicals and dosed chemicals are different"))

# Filtering out non-circulation mediums (these are retained in cvtdb_original)
cvt_df <- cvt_df %>%
  filter(conc_medium_normalized %in% c("blood", "plasma"))

quick_message(text = c("Filtering out:",
                       "Non-plasma/blood medium collected."))
